Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 300\,pc Halo

### 2021-02-09 cluster_paper/ms.tex
update paper

\newcommand{\nautorot}{1{,}144\ } % % depends on which cuts are adopted. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt

@@

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  The
  stars in these ``halos'' could have escaped from the cores of the
  cluster, or they could be remnants of a star formation complex that
  has since dispersed.  They could also be field stars---false
  positives from overly aggressive clustering algorithms.  In this
  study, we focus on a 300\,pc halo reported
  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical
  tidal radius of the cluster is {\bf $\sim$XX\,pc}.  Using photometry
  from Gaia, rotation periods from TESS, and lithium measurements from
  Gaia-ESO and GALAH, we find that the halo of NGC\,2516 is real and
  coeval with the core.  Two in three kinematically selected halo
  members 25-250\,pc from the cluster center show rotation periods
  consistent with a gyrochronological age of 120\,Myr.  Stars in the
  inner 25\,pc show higher period detection rates than the
  halo (73\% vs.\ 66\%), but both vastly exceed our field star
  comparison sample ({\bf XX\%}).  The rotation rate and lithium abundance
  are also anti-correlated at fixed stellar mass, as is the case with
  comparable clusters.  This work expands the set of confirmed
  NGC\,2516 members by a factor of $\approx$2, and quantifies the
  degree to which Gaia-based analyses can identify dispersed stellar
  populations against a background of field star contaminants.  We
  highlight implications for spectroscopic survey targeting, open
  cluster dispersal, and planet searches around young stars.
\end{abstract}
% Rotators < 25pc : 0.72929 +/- 0.00013
% Rotators > 25pc : 0.66172 +/- 0.00023

@@

Outside of the issue of star formation, identifying the remnant halos of
open clusters is important for a separate project: that of discovering
young transiting planets.  Young transiting planets are hard to find
because young stars are rare (CITE), and often reside in crowded regions
of the sky (CITE).  If the halos of nearby star clusters can be reliably
identified, this could expand the census of nearby sub-Gyr stars by a
factor of 2 or 3 (CITE).  A fortuitous benefit of searching for planets
in cluster environments is that issues with stellar crowding are also
alleviated.

Recent clustering studies using Gaia have begun to report the
identification of structures that could correspond to low-density halos
of stars that may have evaporated from open clusters (e.g., Kounkel+19,
Kounkel+20, Meingast+21).  However, different clustering methods on the
Gaia data tend to give different results (CITE: Hunt \& Reffert 2020).
Employing say Gaussian Mixture Modelling, or any analogous method that
requires ``clusters'' to be ellipses in phase-space (position or
velocity) unsurprisingly yields open clusters that are roughly elliptical.
Some unsupervised clustering methods lead to comparable ``roughly
elliptical'' results \citep[][hereafter
\citetalias{cantatgaudin_gaia_2018}]{cantatgaudin_gaia_2018}.  Other
unsupervised methods such as HDBScan (e.g., \citep[][hereafter
\citetalias{kounkel_untangling_2019}]{kounkel_untangling_2019}) have
been found to yield additional structures, particularly in lower density
regions such as the Psc-Eri stream (CITE: Meingast+18, Curtis+19,
Newton+21, but reword bc discovery didn't use HDBScan), but also more
generally around many open clusters \citep{kounkel_untangling_2019}.
Unsupervised approaches that incorporate physical constraints like
imposing a maximum velocity dispersion on putative members yield similar
results, potentially with higher purity (Meingast+21).

@@

% Theia 613
\citetalias{cantatgaudin_gaia_2018} reported 1106 candidate members of
\cn, brighter than $G=18$\,mag.  \citet{kounkel_untangling_2019}
reported 3003 candidate members, and included stars up to
$\approx$1\,mag fainter.  Each study used unsupervised clustering
based on the second Gaia Data Release, detailed in
Appendix~\ref{app:clustering}.

@@

>>>>

%\newcommand{\nautorot}{1{,}144\ } % % For nequal<=0, nclose<=100, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorot}{987\ } % % For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt

@@

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  The
  stars in these ``halos'' could have escaped from the cores of the
  cluster, or they could be remnants of a star formation complex that
  has since dispersed.  They could also be field stars---false
  positives from overly aggressive clustering algorithms.  In this
  study, we focus on a 300\,pc halo reported
  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical
  tidal radius of the cluster is $\sim$10\,pc.  Using photometry
  from Gaia, rotation periods from TESS, and lithium measurements from
  Gaia-ESO and GALAH, we find that the halo of NGC\,2516 is real and
  coeval with the core.  Two in three kinematically selected halo
  members 25-250\,pc from the cluster center show rotation periods
  consistent with a gyrochronological age of 120\,Myr.  Stars in the
  inner 25\,pc show higher period detection rates than the
  halo (73\% vs.\ 66\%), but both vastly exceed our field star
  comparison sample ({\bf XX\%}).  The rotation rate and lithium abundance
  are also anti-correlated at fixed stellar mass, as is the case with
  comparable clusters.  This work expands the set of confirmed
  NGC\,2516 members by a factor of $\approx$2, and quantifies the
  degree to which Gaia-based analyses can identify dispersed stellar
  populations against a background of field star contaminants.  We
  highlight implications for spectroscopic survey targeting, open
  cluster dispersal, and planet searches around young stars.
\end{abstract}
% Rotators < 25pc : 0.72929 +/- 0.00013
% Rotators > 25pc : 0.66172 +/- 0.00023

@@

Outside of the issue of star formation, identifying the remnant halos of
open clusters is important for a separate project: that of discovering
young transiting planets.  Young transiting planets are hard to find
because young stars are rare (CITE), and often reside in crowded regions
of the sky (CITE).  If the halos of nearby star clusters can be reliably
identified, this could expand the census of nearby sub-Gyr stars by a
factor of 2 or 3 (CITE).  
One path to identify these young stars is to combine photometric
selection criteria with spectroscopic observations (e.g., CITE Zerjal
2019, 2020).
A fortuitous benefit of searching for planets
in cluster environments is that issues with stellar crowding are also
alleviated.

Recent clustering studies using Gaia have begun to report the
identification of structures that could correspond to low-density
halos of stars that may have evaporated from open clusters (e.g.,
Kounkel+19, Kounkel+20, Meingast+21).  There is evidence that in some
cases, these halos correspond to tidal tails, e.g., in the Hyades
(CITE Meingast+Alves 20... and Roser+20), and in the Ursa Major moving
group (CITE Gagne+20).  However, different clustering methods on the
Gaia data tend to give different results (CITE: Hunt \& Reffert 2020).
Employing say Gaussian Mixture Modelling, or any analogous method that
requires ``clusters'' to be ellipses in phase-space (position or
velocity) unsurprisingly yields open clusters that are roughly
elliptical.  Some unsupervised clustering methods lead to comparable
``roughly elliptical'' results \citep[][hereafter
\citetalias{cantatgaudin_gaia_2018}]{cantatgaudin_gaia_2018}.  Other
unsupervised methods such as HDBScan (e.g., \citep[][hereafter
\citetalias{kounkel_untangling_2019}]{kounkel_untangling_2019}) have
been found to yield additional structures, particularly in lower
density regions such as the Psc-Eri stream (CITE: Meingast+18,
Curtis+19, Newton+21, but reword bc discovery didn't use HDBScan), but
also more generally around many open clusters
\citep{kounkel_untangling_2019}.  Unsupervised approaches that
incorporate physical constraints like imposing a maximum velocity
dispersion on putative members yield similar results, potentially with
higher purity (Meingast+21).

@@

% Theia 613
\citetalias{cantatgaudin_gaia_2018} reported 1106 candidate members of
\cn, brighter than $G=18$\,mag.  \citet{kounkel_untangling_2019}
reported 3003 candidate members, and included stars up to
$\approx$1\,mag fainter.  Each study used unsupervised clustering
based on the second Gaia Data Release, detailed in
Appendix~\ref{app:clustering}.
%FIXME 
\citetalias{cantatgaudin_gaia_2018} considered a cone of radius {\bf
XX DEGREES} centered on the cluster's position.
Given the classically reported tidal radius of {\it e.g.,} 9\,pc, CITE
Piskunov+2008, which is 8$^\circ$ at the cluster distance of
$\approx$400\,pc, perhaps they should have considered stars further
out.

@@

One relevant connection may be to the ``Mamajek-2'' stellar group
(see Jilinski+2009).
